BlueTracer: \\ a Robust API Tracer for Evasive Malware

### 2018-07-12 background.tex
Joe Section Added

Despite being extremely simple to implement, IAT patching suffers from a couple of disadvantages, which significantly limit its use in practice:
\begin{itemize}
    \item It is incredibly easy to detect by simply examining the entries of the IAT and checking whether or not each address falls inside the memory range of the DLL that should contain the function \cite{HookingDetection}.
    \item It is ineffective when function pointers are acquired dynamically, e.g., via \texttt{LoadLibrary} and \texttt{GetProcAddress} \cite{Buescher:2011:BIS:2186328.2186347}.
\end{itemize}

@@

However, in a similar manner to what occurs for IAT patching, it can be easily detected by simply examining the entries of the EAT and checking whether or not each RVA, when added to the DLL base address, falls within the DLL memory range \cite{Stuttard:2014:ADC:2616217}.

@@

\textit{Inline Hooking}, which was made famous by its employment  in the Microsoft \textit{Detours} Windows API hooking library, is one of the most used API hooking techniques since it offers a number of advantages:
\begin{itemize}
\item It is fast and efficient.
\item It can be utilized to hook any code, not just operating systems APIs, but also programmer defined functions \cite{Rootkit}. 
\item Unlike IAT patching, the type of command used to call the function does not matter, meaning that the hooking will be effective regardless of the fact that a function is called using the IAT or using \texttt{LoadLibrary} together with \texttt{GetProcAddress}.
\end{itemize}

@@

On the other hand, a debugger is much easier to detect. In fact, there exist specific Windows APIs whose purpose is to find out whether or not the current process is being debugged. For example, \texttt{IsDebuggerPresent} allows to determine if the calling process is running under a debugger, while \texttt{CheckRemoteDebuggerPresent} checks for the presence of a debugger in a separate process. 
In addition, the \texttt{INT 3} instruction in an API entry point immediately gives away the debugger's presence \cite{HookingDetection}.

>>>>

Despite being extremely simple to implement, IAT patching suffers from a couple of disadvantages, which significantly limit its use in practice:
\begin{itemize}
    \item It is incredibly easy to detect by simply examining the entries of the IAT and checking whether each address falls inside the memory range of the DLL that should contain the function \cite{HookingDetection}.
    \item It is ineffective when function pointers are acquired dynamically, e.g., via \texttt{LoadLibrary} and \texttt{GetProcAddress} \cite{Buescher:2011:BIS:2186328.2186347}.
\end{itemize}

@@

However, in a similar manner to what occurs for IAT patching, it can be easily detected by simply examining the entries of the EAT and checking whether each RVA, when added to the DLL base address, falls within the DLL memory range \cite{Stuttard:2014:ADC:2616217}.

@@

\textit{Inline Hooking}, which was made famous by its employment  in the Microsoft \textit{Detours} Windows API hooking library, is one of the most used API hooking techniques
\\
\\
since it offers a number of advantages:
\begin{itemize}
\item It is fast and efficient.
\item It can be utilized to hook any code, not just operating systems APIs, but also programmer defined functions \cite{Rootkit}. 
\item Unlike IAT patching, the type of command used to call the function does not matter, meaning that the hooking will be effective regardless of the fact that a function is called using the IAT or using \texttt{LoadLibrary} together with \texttt{GetProcAddress}.
\end{itemize}

@@

On the other hand, a debugger is much easier to detect. In fact, there exist specific Windows APIs whose purpose is to find out whether the current process is being debugged. For example, \texttt{IsDebuggerPresent} allows to determine if the calling process is running under a debugger, while \texttt{CheckRemoteDebuggerPresent} checks for the presence of a debugger in a separate process. 
In addition, the \texttt{INT 3} instruction in an API entry point immediately gives away the debugger's presence \cite{HookingDetection}.

### 2018-07-12 conclusions.tex
Joe Section Added

In order to deal with the astounding amount of malware instances surfacing every day, there has been in recent years a shift towards the use of dynamic malware analysis techniques over static ones. A popular dynamic analysis strategy is function call monitoring, as it enables the analyst to grasp the behavior of the malware sample under analysis, despite the presence of code obfuscations. Such technique is normally implemented via API hooking, that is, the interception of functions exported by dynamically linked libraries (DLL). However, API hooking has to face the growing trend of evasive malware strains, which conceal their damaging behavior when under the magnifying glass of an analyst. Furthermore, most of the current API hooking techniques presented in literature do not provide adequate countermeasures against evasive malware.

@@

Finally, the tests conducted on the Al-Khaser anti-malware system stress tool confirmed BlueTracer's ability to provide extensive and informative logs while, at the same time, remaining undetected.

@@

The obvious next step in BlueTracer's development is to test it with a set of malware samples, in order to evaluate its effectiveness in a real-life scenario.

>>>>

\vspace{-0.6cm}
In order to deal with the astounding amount of malware instances surfacing every day, there has been in recent years a shift towards the use of dynamic malware analysis techniques over static ones. A popular dynamic analysis strategy is function call monitoring, as it enables the analyst to grasp the behavior of the malware sample under analysis, despite the presence of code obfuscations. Such technique is normally implemented via API hooking, that is, the interception of functions exported by dynamically linked libraries (DLL). However, API hooking has to face the growing trend of evasive malware strains, which conceal their damaging behavior when under the magnifying glass of an analyst. Furthermore, most of the current API hooking techniques presented in literature do not provide adequate countermeasures against evasive malware.

@@

Finally, the tests conducted on the Al-Khaser anti-malware system stress tool and on real evasive samples collected by Joe Security confirmed BlueTracer's ability to provide extensive and informative logs while, at the same time, remaining undetected.

@@

The obvious next step in BlueTracer's development is to test it with larger a set of highly-evasive malware samples, in order to further evaluate its effectiveness in a real-life scenario.
